Recalibrating Gravitational Wave Phenomenological Waveform Model

### 2023-05-03 src/tex/ms.tex
Working through result

To evaluate the effectiveness of our optimization approach, we analyze
approximately 530 NR waveforms from the SXS catalog in this study. We
specifically select waveforms with negligible eccentricity
(${e<2\times10^{-3}}$) and precession (${ \chi_{x,y}<5\times10^{-3}}$) that are
consistent with the constraints of the waveform model. As shown in
Fig.~\ref{fig:intrin_space}, the intrinsic parameters of the chosen test
waveforms fall within the parameter space defined by the training waveforms.
This suggests that these test waveforms can be used for direct comparison with
the original model.

We begin by investigating the impact of joint optimization on an individual
waveform. Fig. \ref{fig:0154} illustrates that the optimized waveform
outperforms the original waveform, particularly in the inspiral region, where
the amplitude displays a $50\%$ reduction in error. To ensure fairness, we
selected one of the testing waveforms from the catalog presented in
\citep{khan2016frequency}.

We use the constant PSD weighted loss function, $\mathcal{L}_{\mathrm{mean}}$, to evaluate the
mismatch of all testing waveforms, and present the resulting distribution in
Fig. \ref{fig:q148}. The peak of the distribution has shifted towards a lower
mismatch, with a decrease of almost one order of magnitude and a 50\% reduction
in the median. When using $\mathcal{L}_{\mathrm{fl}}$, we observe a similar
improvement with a 22.9\% decrease in the median. However, the distribution lacks
a clear peak due to the problem discussed in Section \ref{subsec:optimization}.
Despite this, both distributions exhibit significant improvement, and the
optimization scheme remains unaffected.

@@

Applying the same methods, we observe that the distributions of mismatches
calculated using the {\zdethp} PSD exhibit superior improvement
compared to the weighted mismatch. The shape of the distribution is similar to
that shown in Fig. \ref{fig:q148}. This result is expected, as the IMRPhenomD
model was initially constructed and fitted using the {\zdethp} weighted
mismatch. Consequently, the model is anticipated to closely fit the NR waveforms
while incorporating the influence of the {\zdethp} PSD, rather than a
constant PSD.

Motivated by the successful improvement of the waveforms, we expand the training
dataset to optimize additional waveforms listed in Tab. \ref{tab:q1248}. The new
set of coefficients generated from this optimization process yields only
marginal improvements in the newly produced waveforms, as seen in Fig.
\ref{fig:q148_q1248_compare}. The high mismatch tail of the optimized
distribution remains comparable in length and endpoint to the original
distribution, indicating that our procedure is incapable of improving these
waveforms. Similarly, utilizing the {\zdethp} PSD to optimize the loss
function with additional waveforms leads to minimal improvement in the resulting
distribution.

>>>>

To evaluate the effectiveness of the optimization, we analyze approximately 530
NR waveforms \kw{Just say how many you have} from the SXS catalog in this study.
We specifically select waveforms with negligible eccentricity
(${e<2\times10^{-3}}$) and precession (${ \chi_{x,y}<5\times10^{-3}}$) that are
consistent with the constraints of the waveform model. As shown in
Fig.~\ref{fig:intrin_space}, the intrinsic parameters of the chosen test
waveforms fall within the parameter space defined by the training waveforms.
This suggests that these test waveforms can be used for direct comparison with
the original model.

To illustrate the improvement on an individual waveform level, we compare a
particular waveform before and after optimization in Fig. \ref{fig:0154}.
Compared to the original IMRPhenomD waveform, we can see the optimized waveform
has smaller residual from NR waveform both in amplitude and phase, particularly
in the inspiral region, where the amplitude displays a $50\%$ reduction in
error. For a fair comparison, we selected one of the testing waveforms from the
catalog presented in \citep{khan2016frequency}.

To quantify the effect of optimization over the entire dataset, We evaluate the
mismatch of all testing waveforms using a constant PSD weighted loss function,
$\mathcal{L}_{\mathrm{mean}}$. We present the resulting distribution in Fig.
\ref{fig:q148}. The peak of the distribution has shifted towards a lower
mismatch, with a decrease of almost one order of magnitude and a 50\% reduction
in the median. When using $\mathcal{L}_{\mathrm{fl}}$, we observe a similar
improvement with a 22.9\% decrease in the median. Even though the distribution
lacks a clear peak due to the problem discussed in Section
\ref{subsec:optimization}, both distributions show observable improvement, and
the optimization scheme remains unaffected \kw{What about the optimization scheme?}.

@@

\kw{Which plot is this paragraph referring to?}
Applying the same methods, we observe that the distributions of mismatches
calculated using the {\zdethp} PSD exhibit superior improvement compared to the
weighted mismatch \kw{What do you mean by weighted mismatch? Like flat PSD?}.
The shape of the distribution is similar to that shown in Fig. \ref{fig:q148}.
This result is expected, as the IMRPhenomD model was initially constructed and
fitted using the {\zdethp} weighted mismatch. Consequently, the model is
anticipated to closely fit the NR waveforms while incorporating the influence of
the {\zdethp} PSD, rather than a constant PSD.

Motivated by the successful improvement of the waveforms, we expand the training
dataset to optimize additional waveforms listed in Tab. \ref{tab:q1248} \kw{What
special about the q=2 events?}. The new set of coefficients generated from this
optimization process yields only marginal improvements in the newly produced
waveforms, as seen in Fig. \ref{fig:q148_q1248_compare}. The high mismatch tail
of the optimized distribution remains comparable in length and endpoint to the
original distribution, indicating that our procedure is incapable of improving
these waveforms. Similarly, utilizing the {\zdethp} PSD to optimize the loss
function with additional waveforms leads to similar improvement in the resulting
distribution.
